Traffic Aware Resource Allocation Schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} Systems

### 2015-07-16 Report/kkt_mse_approach.tex
updated with nam response

In the proposed scheme, the maximum rate constraints are implicitly handled by the objective function without additional rate constraints. However, due to the non-differential objective, the \ac{KKT} conditions cannot be formulated. The non-differentiability is due to the absolute operator present in the norm function. To make the objective differentiable, we consider the following two cases for which the absolute operator can be ignored without affecting the solution, namely,
\begin{itemize}
\item when the exponent \me{q} is even, or
\item when the number of backlogged packets of each user is large enough, \textit{i.e}, \me{Q_k \gg \sum_{n=1}^N \sum_{l=1}^L {t}_{l,k,n}} to ignore the absolute operator and queues in the first place as well.
\end{itemize}

@@

\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \tfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1}

@@

	\alpha^{(i)}_{l,k,n} =& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \tfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1}

@@

Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in \eqref{kkt-mse-4}, one has to be fixed to optimize for the other. So, \me{\alpha^{(i)}} is fixed to evaluate \me{\sigma^{(i)}} using \eqref{kkt-mse-4}. At each iteration, the dual variables \me{\alpha^{(i)}} are linearly interpolated with any point between the fixed iterate \me{\alpha^{(i-1)}} and \me{\tfrac{\sigma^{(i)}}{\epsilon^{(i)}}} using a step size \me{\rho \in (0,1)}. The choice of \me{\rho} depends on the system model and it affects the convergence behavior. It reduces the oscillations in the objective function when \me{\sigma^{(i)}} is negative due to over-allocation.

@@

To obtain a practical distributed precoder design, we assume that each \ac{BS} \me{b} knows the corresponding equivalent channel \me{\mvec{w}{l,k,n}^\herm \mvec{H}{b,k,n}, \forall k \in {\mc{U}}}, which includes the receivers \me{\mvec{w}{l,k,n}}, through precoded uplink pilot signaling. We extend the decentralization methods discussed in \cite{komulainen2013effective}, for the current problem as follows. After receiving the updated transmit precoders from all \acp{BS} in \me{\mathcal{B}}, each user evaluates the \ac{MMSE} receiver in \eqref{kkt-mse-4.6} and notify them to the \acp{BS} via uplink precoded pilots. On receiving pilot signals, \acp{BS} update the \ac{MSE} in \eqref{mse-error} as 
\begin{equation}
\epsilon_{l,k,n}^{(i)} = 1 - \mvec{w}{l,k,n}^{(i)\herm} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)}.
\end{equation}
Using the current \ac{MSE} value, \me{t_{l,k,n}^{(i)},\sigma_{l,k,n}^{(i)}}, and \me{\alpha_{l,k,n}^{(i)}} are evaluated using \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1}, and the updated dual variables \me{\alpha_{l,k,n}} are exchanged between the \acp{BS} to evaluate the transmit precoders \me{\mvec{m}{l,k,n}^{(i+1)}} for the next iteration. The \ac{SCA} operating point is also updated with the current \ac{MSE} value.
\begin{algorithm}
	\SetAlgoLined
	\DontPrintSemicolon
	\BlankLine
	\SetKwInput{KwInit}{Initialize}
	\KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
	\KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
	\KwInit{\me{i=1}, \me{\mbf{w}^{(0)}_{l,k,n}, \tilde{\epsilon}_{l,k,n}} randomly, dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}, and \me{I_{\max}} for certain value}
	\ForEach{\ac{BS} \me{b \in \mc{B}}}{
		\Repeat{until convergence or \me{i \geq I_{\max}}}{
			update \me{\mvec{m}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.3}, and perform precoded downlink pilot transmission \;
			find \me{\mvec{w}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.6} at each user \;
			evaluate \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at each user with the updated \me{\mvec{w}{l,k,n}^{(i)}} \;
			using precoded uplink pilots, \me{\mvec{m}{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} are notified to all \acp{BS} in \me{\mathcal{B}} \;
			\me{i = i + 1} \;
		}
	}
	\caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
	\label{algo-4}
\end{algorithm}

To avoid the back-haul exchanges between the \acp{BS}, as an alternative approach, users perform all processing required and \acp{BS} will update the precoders based on the feedback information from the users. Upon receiving the transmit precoders from \acp{BS}, each user will update the receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{MSE} \me{\epsilon_{l,k,n}}, and the dual variables \me{\sigma_{l,k,n}} and \me{\alpha_{l,k,n}}. The updated \me{\alpha_{l,k,n}} and \me{\mvec{w}{l,k,n}} are notified to the \acp{BS} using two separate precoded uplink pilot symbols with \me{\tilde{\mbf{w}}_{l,k,n}^{(i)} = \sqrt{\alpha_{l,k,n}^{(i)}}\mvec{w}{l,k,n}^{\ast(i)}} and \me{\bar{\mbf{w}}_{l,k,n}^{(i)} = \alpha_{l,k,n}^{(i)}\mvec{w}{l,k,n}^{\ast(i)}} as the precoders. After receiving the precoded uplink pilots, each \ac{BS} use the effective channel \me{\mvec{H}{b,k,n}^{\mathrm{T}} \tilde{\mbf{w}}_{l,k,n}^{(i)}} and \me{\mvec{H}{b,k,n}^{\mathrm{T}} \bar{\mbf{w}}_{l,k,n}^{(i)}} in \eqref{kkt-mse-4.3} to update the transmit precoders, where \me{\mbf{x}^{\ast}} is the complex conjugate of \me{\mbf{x}}. Algorithm \ref{algo-4} outlines a practical way of updating the transmit and the receive beamformers by using \ac{OTA} signaling with the precoded pilots for the \ac{KKT} based \ac{MSE} reformulated \ac{JSFRA} problem. The convergence of the algorithm is discussed in Appendix \ref{sec-dist-conv}.

>>>>

In the proposed scheme, the maximum rate constraints are implicitly handled by the objective function without any need for explicit constraints. However, the \ac{KKT} conditions cannot be formulated due to the non-differentiability of the absolute value operator present in the objective function. In order to make the objective function differentiable, we consider the following two cases for which the absolute operator can be ignored without affecting the optimal solution, namely,
\begin{itemize}
\item when the exponent \me{q} is even, or
\item when the number of backlogged packets of each user is large enough, \textit{i.e}, \me{Q_k \gg \sum_{n=1}^N \sum_{l=1}^L {t}_{l,k,n}} to ignore the absolute operator and queues in the first place as well.
\end{itemize}

@@

\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho^{(i)} \left ( \tfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1}

@@

	\alpha^{(i)}_{l,k,n} =& \alpha^{(i-1)}_{l,k,n} + \rho^{(i)} \left ( \tfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1}

@@

Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in \eqref{kkt-mse-4}, one has to be fixed to optimize for the other. So, \me{\alpha^{(i)}} is fixed to evaluate \me{\sigma^{(i)}} using \eqref{kkt-mse-4}. At each iteration, the dual variables \me{\alpha^{(i)}} are linearly interpolated with any point between the fixed iterate \me{\alpha^{(i-1)}} and \me{\tfrac{\sigma^{(i)}}{\epsilon^{(i)}}} using a diminishing or a fixed step size \me{\rho^{(i)} \in (0,1)}. The choice of \eqn{\rho^{(i)}} affects the convergence behavior and also controls the oscillations in the users rate when \me{\sigma^{(i)}} is negative due to over-allocation.

@@

To obtain a practical distributed precoder design, we assume that each \ac{BS} \me{b} knows the corresponding equivalent channels \me{\mvec{w}{l,k,n}^\herm \mvec{H}{b,k,n}, \forall k \in {\mc{U}}}, which includes the receivers \me{\mvec{w}{l,k,n}}, through precoded uplink pilot signaling. We extend the decentralization methods discussed in \cite{komulainen2013effective}, for the current problem as follows. Upon receiving the updated transmit precoders from all \acp{BS} in \me{\mathcal{B}}, each user will evaluate the \ac{MMSE} receiver \eqref{kkt-mse-4.6} and notify to all \acp{BS} by using precoded uplink pilots. On receiving the pilots, each \ac{BS} update the \ac{MSE} in \eqref{mse-error} as 
\begin{equation}
\epsilon_{l,k,n}^{(i)} = 1 - \mvec{w}{l,k,n}^{(i)\herm} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)}.
\end{equation}
Using the current \ac{MSE} value, \me{t_{l,k,n}^{(i)},\sigma_{l,k,n}^{(i)}}, and \me{\alpha_{l,k,n}^{(i)}} are evaluated using \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1}, and the updated dual variables \me{\alpha_{l,k,n}} are exchanged between the \acp{BS} to evaluate the transmit precoders \me{\mvec{m}{l,k,n}^{(i+1)}} for the next iteration. The \ac{SCA} operating point is also updated with the current \ac{MSE} value.
\begin{algorithm}
	\SetAlgoLined
	\DontPrintSemicolon
	\BlankLine
	\SetKwInput{KwInit}{Initialize}
	\KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
	\KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
	\KwInit{\me{i=1}, \me{\mbf{w}^{(0)}_{l,k,n}, \tilde{\epsilon}_{l,k,n}} randomly, dual variables \me{{\alpha}_{l,k,n}^{(0)} = 1}, and \me{I_{\max}} for certain value}
	\ForEach{\ac{BS} \me{b \in \mc{B}}}{
		\Repeat{until convergence or \me{i \geq I_{\max}}}{
			update \me{\mvec{m}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.3}, and perform precoded downlink pilot transmission \;
			find \me{\mvec{w}{l,k,n}^{(i)}} using \eqref{kkt-mse-4.6} at each user \;
			evaluate \me{\epsilon_{l,k,n}^{(i)}}, \me{t_{l,k,n}^{(i)}}, \me{\sigma_{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} using \eqref{kkt-mse-4.4} and \eqref{kkt-mse-4.5}, \eqref{kkt-mse-4.2} and \eqref{kkt-mse-4.1} at each user with the updated \me{\mvec{w}{l,k,n}^{(i)}} \;
			using precoded uplink pilots, \me{\mvec{m}{l,k,n}^{(i)}} and \me{\alpha_{l,k,n}^{(i)}} are notified to all \acp{BS} in \me{\mathcal{B}} \;
			\me{i = i + 1} \;
		}
	}
	\caption{\ac{KKT} approach for the \ac{JSFRA} scheme}
	\label{algo-4}
\end{algorithm}

To avoid the back-haul exchanges between the \acp{BS}, as an alternative approach, users can perform all the processing required and \acp{BS} will update the precoders based on the feedback from all the users. Upon receiving the transmit precoders from the \acp{BS}, each user will update the receive beamformer \me{\mvec{w}{l,k,n}}, the \ac{MSE} \me{\epsilon_{l,k,n}}, and the dual variables \me{\sigma_{l,k,n}} and \me{\alpha_{l,k,n}}. Then, the updated \me{\alpha_{l,k,n}} and \me{\mvec{w}{l,k,n}} are notified to the \acp{BS} using two separate precoded uplink pilots with \me{\tilde{\mbf{w}}_{l,k,n}^{(i)} = \sqrt{\alpha_{l,k,n}^{(i)}}\mvec{w}{l,k,n}^{\ast(i)}} and \me{\bar{\mbf{w}}_{l,k,n}^{(i)} = \alpha_{l,k,n}^{(i)}\mvec{w}{l,k,n}^{\ast(i)}} as the precoders. Upon receiving the precoded uplink pilots, each \ac{BS} evaluate the effective channels \me{\mvec{H}{b,k,n}^{\mathrm{T}} \tilde{\mbf{w}}_{l,k,n}^{(i)}} and \me{\mvec{H}{b,k,n}^{\mathrm{T}} \bar{\mbf{w}}_{l,k,n}^{(i)}} to update the transmit precoders using \eqref{kkt-mse-4.3}, where \me{\mbf{x}^{\ast}} is the complex conjugate of \me{\mbf{x}}. Algorithm \ref{algo-4} outlines a practical way of updating the transmit and the receive beamformers by using \ac{OTA} signaling with the precoded pilots for the \ac{KKT} based \ac{MSE} reformulated \ac{JSFRA} problem. The convergence of the algorithm is discussed in Appendix \ref{sec-dist-conv}.

### 2015-07-16 Report/mse_formulation.tex
updated with nam response

Let \me{{v}^{\prime}_k = Q_k - \sum_{n=1}^N \sum_{l=1}^L t_{l,k,n}} denote the queue deviation corresponding to user \me{k} and \me{\tilde{{v}}^{\prime}_k \triangleq a_k^{1/{q}}v^{\prime}_k} represents the weighted equivalent. By using the relaxed \ac{MSE} expression in \eqref{mse-error}, the problem in \eqref{eqn-3} can be expressed as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{rCl}\label{eqn-mse-1} \neqsub \allowdisplaybreaks
\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n},\\ \epsilon_{l,k,n},\mvec{w}{l,k,n}}} {\text{minimize}} & \quad & \|  \tilde{\mbf{v}}^{\prime}  \|_q \IEEEyessubnumber \label{eqn-mse-1.1} \\
\text{subject to} & \quad & t_{l,k,n} \leq -\log_2(\epsilon_{l,k,n}) \IEEEyessubnumber \label{eqn-mse-1.2} \\
 & \quad & \epsilon_{l,k,n} \geq  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{\mathclap{(j,i) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \right |^2 + \enoise \eqspace \IEEEyessubnumber \label{eqn-mse-1.3} \\
& \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \; \fall b. \IEEEyessubnumber  \eqspace \label{eqn-mse-1.4}
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CL}\label{eqn-mse-1} \neqsub \allowdisplaybreaks
	\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n},\\ \epsilon_{l,k,n},\mvec{w}{l,k,n}}} {\text{minimize}} & \quad \|  \tilde{\mbf{v}}^{\prime}  \|_q \IEEEyessubnumber \label{eqn-mse-1.1} \\
	\text{subject to} & t_{l,k,n} \leq -\log_2(\epsilon_{l,k,n}) \IEEEyessubnumber \label{eqn-mse-1.2} \\
	& \sum_{{(j,i) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \right |^2 + \enoise \nonumber \\	
	& \qquad + \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 \leq \epsilon_{l,k,n}\IEEEyessubnumber \label{eqn-mse-1.3} \\
	& \quad \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \; \fall b. \IEEEyessubnumber  \eqspace \label{eqn-mse-1.4}
\end{IEEEeqnarray}}
\begin{algorithm}
	\SetAlgoLined
	\DontPrintSemicolon
	\BlankLine
	\SetKwInput{KwInit}{Initialize}
	\KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
	\KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
	\KwInit{\me{i=0,j=0} and \me{\tilde{\mbf{m}}_{l,k,n}} using single user beamformer satisfying power constraint \eqref{eqn-9.1d}}
	update \me{\tilde{\mbf{w}}_{l,k,n}, \tilde{\mbf{\beta}}_{l,k,n}} using \eqref{eqn-10} and \eqref{eqn-9.1c} with \me{\tilde{\mbf{m}}_{l,k,n}} \;
	\Repeat{Queue convergence or \me{i \geq I_{\max}}}{
		\Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
			solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9}\;
			update the constraint set \eqref{eqn-8} with \me{{\mbf{u}}_{l,k,n}} and \me{\mvec{m}{l,k,n}} using \eqref{eqn-wsrm-expr}\;
			$j = j + 1$\;
		}
		update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
		$i = i + 1$, \me{j = 0}\;
	}
	\caption{Algorithm of \acs{JSFRA} scheme}
	\label{algo-1} 
\end{algorithm}

>>>>

Let \me{{v}^{\prime}_k = Q_k - \sum_{n=1}^N \sum_{l=1}^L t_{l,k,n}} denote the queue deviation corresponding to user \me{k} and \me{\tilde{{v}}^{\prime}_k \triangleq a_k^{1/{q}}v^{\prime}_k} represents the weighted equivalent. By using the relaxed \ac{MSE} expression in \eqref{mse-error}, the problem in \eqref{eqn-3} can be expressed as
\iftoggle{single_column}{
\begin{IEEEeqnarray}{rCl}\label{eqn-mse-1} \neqsub \allowdisplaybreaks
\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n},\\ \epsilon_{l,k,n},\mvec{w}{l,k,n}}} {\text{minimize}} & \quad & \|  \tilde{\mbf{v}}^{\prime}  \|_q \IEEEyessubnumber \label{eqn-mse-1.1} \\
\text{subject to} & \quad & t_{l,k,n} \leq -\log_2(\epsilon_{l,k,n}) \IEEEyessubnumber \label{eqn-mse-1.2} \\
 & \quad & \epsilon_{l,k,n} \geq  \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 + \sum_{\mathclap{(j,i) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \right |^2 + \enoise \eqspace \IEEEyessubnumber \label{eqn-mse-1.3} \\
& \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \; \fall b. \IEEEyessubnumber  \eqspace \label{eqn-mse-1.4}
\end{IEEEeqnarray}
}{\allowdisplaybreaks
\begin{IEEEeqnarray}{CL}\label{eqn-mse-1} \neqsub \allowdisplaybreaks
	\underset{\substack{t_{l,k,n},\mvec{m}{l,k,n},\\ \epsilon_{l,k,n},\mvec{w}{l,k,n}}} {\text{minimize}} & \quad \|  \tilde{\mbf{v}}^{\prime}  \|_q \IEEEyessubnumber \label{eqn-mse-1.1} \\
	\text{subject to} & t_{l,k,n} \leq -\log_2(\epsilon_{l,k,n}) \IEEEyessubnumber \label{eqn-mse-1.2} \\
	& \sum_{{(j,i) \neq (l,k)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,k,n} \mvec{m}{j,i,n} \right |^2 + \enoise \nonumber \\	
	& \qquad + \left | 1 - \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right |^2 \leq \epsilon_{l,k,n}\IEEEyessubnumber \label{eqn-mse-1.3} \\
	& \quad \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \sum_{l=1}^L \trace \, (\mvec{m}{l,k,n} \mvec{m}{l,k,n}^\herm) \leq P_{{\max}} \; \fall b. \IEEEyessubnumber  \eqspace \label{eqn-mse-1.4}
\end{IEEEeqnarray}}
\begin{algorithm}
	\SetAlgoLined
	\DontPrintSemicolon
	\BlankLine
	\SetKwInput{KwInit}{Initialize}
	\KwIn{\me{a_k, \, Q_k, \, \mvec{H}{b,k,n},\; \fall b \in \mathcal{B}, \, \fall k \in \mathcal{U}, \fall n \in \mathcal{N}}}
	\KwOut{\me{\mvec{m}{l,k,n}} and \me{\mvec{w}{l,k,n} \fall l \in \set{1,2,\dotsc,L}}}
	\KwInit{\me{i=0,j=0} and \me{\tilde{\mbf{m}}_{l,k,n}} using single user beamformer satisfying power constraint \eqref{eqn-9.1d}}
	update \me{\tilde{\mbf{w}}_{l,k,n}, \tilde{\mbf{\beta}}_{l,k,n}} using \eqref{eqn-10} and \eqref{eqn-9.1c} with \me{\tilde{\mbf{m}}_{l,k,n}} \;
	\Repeat{Queue convergence or \me{i \geq I_{\max}}}{
		\Repeat{\ac{SCA} convergence or \me{j \geq J_{\max}}}{
			solve for the transmit precoders \me{\mvec{m}{l,k,n}} using \eqref{eqn-9}\;
			update \me{\tilde{\mbf{u}}_{l,k,n}} and the constraint set \eqref{eqn-8} with the newly found \me{\mvec{m}{l,k,n}} and \me{\mvec{\beta}{l,k,n}} using \eqref{eqn-wsrm-expr}\;
			$j = j + 1$\;
		}
		update the receive beamformers \me{\mvec{w}{l,k,n}} using \eqref{eqn-9--1} or \eqref{eqn-10} with the updated precoders \me{\mvec{m}{l,k,n}}\;
		$i = i + 1$, \me{j = 0}\;
	}
	\caption{Algorithm of \acs{JSFRA} scheme}
	\label{algo-1} 
\end{algorithm}

### 2015-07-16 Report/problem_formulation.tex
updated with nam response

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by limiting it to a constant value. \review{In fact, using \me{\ell_{q=2}} in \eqref{eqn-4.2}, we obtain the following objective
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2
\end{equation}
which is similar to the objective in \eqref{q_gen_sum}. It is achieved either by removing \me{t_k^2} from \eqref{pf-1} or when the total number of queued packets is large for all users such that \me{t_k^2} has no impact on the objective function.}

By limiting \me{t_k^2} with a constant value, the \ac{Q-WSRM} formulation requires an explicit rate constraint \eqref{eqn-3.1.4} to avoid over-allocation of the available resources. In the proposed queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function \eqref{eqn-4.2} itself. It makes the problem simpler and allows us to employ efficient algorithms to distribute the precoder design problem across each \ac{BS} independently with minimal information exchange \cite{boyd2011distributed}. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRME} problems handle the sub-channels jointly to obtain an efficient allocation by identifying the optimal space-frequency resources for the users.

>>>>

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by limiting it to a constant value. In fact, by using \me{\ell_{q=2}} norm in \eqref{eqn-4.2}, we obtain the objective function, similar to \eqref{q_gen_sum} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2
\end{equation}
The equivalence is achieved by either removing \me{t_k^2} from \eqref{pf-1} or when the number of queued packets is large enough.

By ignoring \me{t_k^2} from \eqref{pf-1}, the \ac{Q-WSRM} scheme requires an explicit rate constraint \eqref{eqn-3.1.4} to avoid over-allocation of the resources. In the proposed queue deviation approach, explicit rate constraints are not needed, since they are handled by the objective function \eqref{eqn-4.2} itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRME} problems handle the sub-channels jointly to obtain an efficient allocation by identifying the optimal space-frequency resources for the users.

### 2015-07-16 Report/queue_minimizing_intro.tex
updated - final - draft
updated
updated with nam response

% To begin with, we discuss the existing solution and its extension to solve the problem of resource allocation over multiple sub-channels. Then, we provide two approaches to solve the \ac{JSFRA} precoder design problem to minimize the total number of backlogged packets. At the end, we discuss a low complex solution for the \ac{JSFRA} problem by limiting the precoder design over each sub-channel at a given instant.

\review{In general, the precoder design for the \ac{MIMO} \ac{OFDM} problem is difficult due to its nonconvex nature}. In addition, the objective of minimizing the number of the queued packets over space-frequency dimensions adds further complexity. Since the scheduling of users in each sub-channel attained by allocating zero transmit power over certain sub-channels, our solutions perform joint precoder design and user scheduling. Before discussing the proposed solutions, we consider the existing algorithm to minimize the number of backlogged packets with additional constraints required by the problem.

>>>>

% To begin with, we discuss the existing solution and its extension to solve the problem of resource allocation over multiple sub-channels. Then, we provide two approaches to solve the \ac{JSFRA} precoder design problem to minimize the total number of backlogged packets. At the end, we discuss a low complex solution for the \ac{JSFRA} problem by limiting the precoder design over each sub-channel at a given instant.
In general, the precoder design for the \ac{MIMO} \ac{OFDM} problem is difficult due to its nonconvex nature. In addition, the objective of minimizing the number of the queued packets over space-frequency dimensions adds further complexity. Since the scheduling of users in each sub-channel is achieved by allocating zero transmit power over certain sub-channels, our solutions perform joint precoder design and user scheduling. Before discussing the proposed solutions, we consider the existing algorithm to minimize the number of backlogged packets with additional constraints required by the problem.

### 2015-07-16 Report/sub_channel_wsrm.tex
updated with nam response

As an alternative sub-optimal solution, we present a \acl{QM} \ac{SRA}, which solves for the precoders using \ac{JSFRA} formulation for a specific sub-channel \me{i} with a fixed transmit power \me{P_{\max,i}}. The sharing of power can be either equal or based on predetermined pattern as in partial frequency reuse with \eqn{\sum_{i=1}^N P_{\max,i} = P_{\max}}. Even though \me{N} sub-channels are present at any given scheduling instant, precoders are computed for each sub-channel in a sequential manner with the sub-channel specific total power constraint \me{P_{\max,i}} and the number of backlogged packets. Let \me{Q_{k,i}} be the number of backlogged packets associated with user \me{k} before solving for the precoders specific to the sub-channel \me{i}. Since the precoder design is sequential, \textit{i.e}, the precoders are designed for sub-channels \me{[0,i-1]} before \me{\ith{i}} sub-channel, the number of backlogged packets for the initial sub-channel is initialized as \me{Q_{k,1} = Q_k}. The queues associated with the consecutive sub-channels are given by
\begin{equation}	\label{eqn-weight}
	Q_{k,i+1} = \max \Big ( Q_k - \sum_{j = 1}^{i} \, \sum_{l = 1}^{L} \, t_{l,k,j} ,0 \Big ) \; \forall \; k \in \mathcal{U}
\end{equation}
where \me{t_{l,k,j}} is the \me{\ith{k}} user rate on the sub-channel \me{j}.

>>>>

As an alternative sub-optimal solution, we present a \acl{QM} \ac{SRA}, which solves for the precoders using \ac{JSFRA} formulation for a specific sub-channel \me{i} with a fixed transmit power \me{P_{\max,i}}. The power sharing can either be equal or based on some predetermined limits on each sub-channel as in partial frequency reuse as
\begin{equation}
\sum_{i=1}^N P_{\max,i} = P_{\max}.
\end{equation}
Even though \me{N} sub-channels are present at any given scheduling instant, precoders are computed for each sub-channel sequentially with \me{P_{\max,i}} and the residual number of backlogged packets. Let \me{Q_{k,i}} be the number of backlogged packets associated with user \me{k} while determining the precoders for the \eqn{\ith{i}} sub-channel. Since the precoder design is sequential, \textit{i.e}, the precoders are designed for sub-channels \me{[0,i-1]} before \me{\ith{i}} sub-channel, the number of backlogged packets for the initial sub-channel is initialized as \me{Q_{k,1} = Q_k}. The queues associated with the consecutive sub-channels are given by
\begin{equation}	\label{eqn-weight}
	Q_{k,i+1} = \max \Big ( Q_k - \sum_{j = 1}^{i} \, \sum_{l = 1}^{L} \, t_{l,k,j} ,0 \Big ) \; \forall \; k \in \mathcal{U}
\end{equation}
where \me{t_{l,k,j}} is the \me{\ith{k}} user rate on the sub-channel \me{j}.
